---
title: "case_study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# THIS IS OUR MODEL
$$
T_i \sim Weibull(\alpha, \mu) \\
log(\mu) = \beta_0 + \beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}\\
$$
Log of mu is done so that the mu parameter is always positive (the support on mu in Weibull is that mu > 0)

# MORE SPECIFICALLY
$$
T_i \sim Weibull(\alpha, \mu_i) \\
\log(\mu_i) = \beta_0+\beta_1 \text{(Year of Birth_i)}+\beta_2 \text{(Leadership_i = UsPres)} + \beta_3\text{(Leadership_i = ChinaEmp)} \\
+ \beta_4\text{(Leadership_i = DalaiLama)} +\beta_5\text{(Leadership_i = JapanEmp)}
$$

# https://amy-herring.github.io/STA440/decks/survival.html#/proportional-hazards-models
# generate predictive samples of Obama and Naruhito, then do Pr(B0+B1 > B1+B2)
# < 1.02 Rhat and > 200 n.eff
# bayes plot package

# Libraries
```{r}
library(tidyverse)
library(lubridate)
```

# Data
```{r}
leaders_data <- get(load("Leaders.RData"))
```

# EDA

```{r EDA}
leaders_data$Censored <- as.factor(leaders_data$Censored)
leaders_data$Type <- as.factor(leaders_data$Type)
leaders_data$Age.Event <- as.numeric(leaders_data$Age.Event)
summary(leaders_data)

# boxplot of age (Age.Event) by Type
ggplot(data = leaders_data) +
  geom_boxplot(aes(x = Type, y = Age.Event))

ggplot(data = leaders_data) +
  geom_point(aes(x = Birth.Date, y = Age.Event))

```

# THIS IS OUR MODEL
$$
T_i \sim Weibull(\alpha, \mu) \\
log(\mu) = \beta_0 + \beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}\\
$$
Log of mu is done so that the mu parameter is always positive (the support on mu in Weibull is that mu > 0)

# MORE SPECIFICALLY
$$
T_i \sim Weibull(\alpha, \mu_i) \\
\log(\mu_i) = \beta_0+\beta_1 \text{(Year of Birth_i)}+\beta_2 \text{(Leadership_i = UsPres)} + \beta_3\text{(Leadership_i = ChinaEmp)} \\
+ \beta_4\text{(Leadership_i = DalaiLama)} +\beta_5\text{(Leadership_i = JapanEmp)}
$$

# Data Cleaning

```{r}
# filter out the 4 leaders we need to predict

leaders_data <- leaders_data %>%
  mutate(TypeChinaEmp = as.factor(case_when(Type == "ChinaEmp" ~ 1,
                             TRUE ~ 0)),
         TypeDalaiLama = as.factor(case_when(Type == "DalaiLama" ~ 1,
                                   TRUE ~ 0)),
         TypeJapanEmp = as.factor(case_when(Type == "JapanEmp" ~ 1, 
                                  TRUE ~ 0)),
         TypePope = as.factor(case_when(Type == "Pope" ~ 1, 
                              TRUE ~ 0)),
         TypeUsPres = as.factor(case_when(Type == "UsPres" ~ 1,
                                TRUE ~ 0)))

leaders_nopred <- leaders_data %>%
  filter(!(Name %in% c("14th Dalai Lama (Tenzin Gyatso)", "Naruhito (Emperor Reiwa)", "Barack Obama", "Francis")))

leaders_nopred$Birth.Year <- year(leaders_nopred$Birth.Date)

survival_raw <- leaders_nopred$Age.Event
n <- length(survival_raw)

censored <- leaders_nopred$Censored
survival <- ifelse(censored == 0, survival_raw, NA)

censoring_limits <- ifelse(censored == 0, 120, survival_raw)
x_1 <- leaders_nopred$Birth.Year - mean(leaders_nopred$Birth.Year)
x_2 <- leaders_nopred$TypeUsPres
x_3 <- leaders_nopred$TypeChinaEmp
x_4 <- leaders_nopred$TypeDalaiLama
x_5 <- leaders_nopred$TypeJapanEmp

```

# First Approach Function

```{r Censor-1st-approach-1}
first_approach <- function(){
  for(i in 1:n){
    #
    # Likelihood part
    #
    censored[i] ~ dinterval(survival[i], censoring_limits[i]) 

    # To deal with censoring: 
    # censored equals 1 for Benedict XVI as of 25/12/2016,
    # 0 for the other popes.
    # survival equals NA for Benedict XVI,
    # and the survival times of the other popes.
    # censoring_limits equals 11.7 for Benedict XVI,
    # and values (e.g. 32) greater than or equal to 
    # the survival times of the other popes.
    
    survival[i] ~ dweib(r, mu[i]) # Basic Weibull assumption
    mu[i] <- exp(beta[i])  # Defining beta as log(mu)
    beta[i] <- beta_0 + beta_1*x_1[i] + beta_2*x_2[i] + beta_3*x_3[i] + beta_4*x_4[i] + beta_5*x_5[i]
    
    # beta = log(mu) is a linear function of the covariates
  }
  
  
  # Priors
  beta_0 ~ dnorm(0.0, 1.0E-4) # Prior on beta_0 is normal with low precision
  beta_1 ~ dnorm(0.0, 1.0E-4) # Prior on beta_1 is normal with low precision
  beta_2 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  beta_3 ~ dnorm(0.0, 1.0E-4) # Prior on beta_1 is normal with low precision
  beta_4 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  beta_5 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  r ~ dexp(0.001) # Prior on r
  
  # Define the alphas
  alpha_0 <- - beta_0 / r
  alpha_1 <- - beta_1 / r
  alpha_2 <- - beta_2 / r
  alpha_3 <- - beta_3 / r
  alpha_4 <- - beta_4 / r
  alpha_5 <- - beta_5 / r
  

  # Percentage increases
  # percentage_increase_age <- 100*(exp(alpha_1) - 1)
  # percentage_increase_year <- 100*(exp(alpha_2) - 1)
  
  # Posterior median at various covariate values
  # beta_med <- beta_0 + beta_1*x_1_new + beta_2*x_2 + beta_3*x_3 + beta_4*x_4 + beta_5*x_5 
  # # New values need to be supplied
  # t_median <- pow(log(2) * exp(-beta_med), 1 / r)
  
  # Predictive distribution of age at the new values
  beta_Francis <- beta_0 + beta_1*year_Francis 
  # Values of age_Francis and year_Francis need to be provided
  mu_Francis <- exp(beta_Francis)
  survival_Francis ~ dweib(r, mu_Francis) %_% T(present_length, upper_length) 
  # Take into account the current pontificate length
  # Also specify a sensible upper bound
  age_Francis_predictive <- survival_Francis 
  # Work also with age
}
```

```{r 1st-approach-Francis-pred, warning=F, message=F}
library(R2jags)
# Prepare the data
age_new <- 60
x_1_new <-  age_new - mean(leaders_nopred$Birth.Year)
# year_new <- 1750
# x_2_new <- year_new - mean(popes$Year.Elected)

# Data specific to Francis
birth_year_Francis <- 1936
year_Francis <- birth_year_Francis - mean(leaders_nopred$Birth.Year)

# Predictive distribution of pontificate length for Francis at election
# Conditional on being greater than 7.46 years (Aug 24, 2020)
present_length <- leaders_data$Age.Event[leaders_data$Name =="Francis"]
upper_length <- 120
data_Popes_first_approach <- list("n",
                   "censored",
                   "survival",
                   "censoring_limits",
                   "x_1",
                   "x_2",
                   "x_3",
                   "x_4",
                   "x_5",
                   "x_1_new",
                   #"x_2_new",
                   #"age_at_election",
                   #"age_Francis",
                   "year_Francis",
                   "present_length",
                   "upper_length")
Bayesian_Popes_first_approach <- jags(data = data_Popes_first_approach,  
                       parameters.to.save = c("beta_0",
                                              "beta_1",
                                              "beta_2",
                                              "beta_3",
                                              "beta_4",
                                              "beta_5",
                                              "alpha_0",
                                              "alpha_1",
                                              "alpha_2",
                                              "alpha_3",
                                              "alpha_4",
                                              "alpha_5",
                                              #"percentage_increase_age",
                                              #"percentage_increase_year",
                                              "r",
                                              #"t_median",
                                              "survival_Francis",
                                              "age_Francis_predictive"), 
                       n.iter = 50000, 
                       n.chains = 3,
                       model.file = first_approach)

Bayesian_Popes_first_approach
```

