---
title: "case_study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# THIS IS OUR MODEL
$$
T_i \sim Weibull(\alpha, \mu) \\
log(\mu) = \beta_0 + \beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}\\
$$
Log of mu is done so that the mu parameter is always positive (the support on mu in Weibull is that mu > 0)

# MORE SPECIFICALLY
$$
T_i \sim Weibull(\alpha, \mu_i) \\
\log(\mu_i) = \beta_0+\beta_1 \text{(Year of Birth_i)}+\beta_2 \text{(Leadership_i = UsPres)} + \beta_3\text{(Leadership_i = ChinaEmp)} \\
+ \beta_4\text{(Leadership_i = DalaiLama)} +\beta_5\text{(Leadership_i = JapanEmp)}
$$

According to Graham, the following is being done by the model above. So basically Stander et al is doing cox proportional, but they just didn't write out this hazard function explicity in the paper.
$$
h_i(t)=h_0(t)\exp(\beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}) \\
$$
hazard of death is a rate -- instantaneous failure rate (probability is integral from 0 to J over it dt)

$$
f(t) = \alpha\lambda t^{\alpha-1}\exp(-\lambda t^\alpha)
$$
run JAGS mcmc on the betas, follow Stander et al code as a guide




# EXPERIMENTAL CODE
```{r}
leaders_data <- get(load("Leaders.RData"))
```

```{r}
library(dynsurv)
library(survival)
library(tidyverse)
library(lubridate)
data(tooth)
```


```{r}
library(tidyverse)
mydata = leaders_data
mygrid <- tooth.grid # Amy said the time variable is not relevant
myformula <- Surv(leaders_data$Birth.Date, leaders_data$Event.Date, type = "interval2") ~ leaders_data$Birth.Date + leaders_data$Type
```

# https://amy-herring.github.io/STA440/decks/survival.html#/proportional-hazards-models
# generate predictive samples of Obama and Naruhito, then do Pr(B0+B1 > B1+B2)
# < 1.02 Rhat and > 200 n.eff
# bayes plot package

```{r}
fit <- bayesCox(myformula, mydata, mygrid, 
                 out = "leadersCox.txt", 
                 model = "TimeIndep", 
                 base.prior = list(type = "Gamma", shape = 0.1, rate = 0.1),    
                 coef.prior = list(type = "Normal", mean = 0, sd = 1),  
                 gibbs = list(iter = 100, burn = 20, thin = 1, verbose = TRUE, nReport = 5))
plotCoef(coef(fit, level = 0.9))
```
```{r}
# https://www.rdocumentation.org/packages/dynsurv/versions/0.3-3/topics/bayesCox
mydata0 <- tooth
mygrid0 = tooth.grid
myformula0<- Surv(left, right, type = "interval2") ~ dmf + sex
```

```{r}
fit0 <- bayesCox(myformula0, mydata0, mygrid0, out = "tiCox.txt",
                 model = "TimeIndep",
                 base.prior = list(type = "Gamma", shape = 0.1, rate = 0.1),
                 coef.prior = list(type = "Normal", mean = 0, sd = 1),
                 gibbs = list(iter = 100, burn = 20, thin = 1, verbose = TRUE, nReport = 5))
plotCoef(coef(fit0, level = 0.9))
```

# EDA

```{r EDA}
leaders_data$Censored <- as.factor(leaders_data$Censored)
leaders_data$Type <- as.factor(leaders_data$Type)
leaders_data$Age.Event <- as.numeric(leaders_data$Age.Event)
summary(leaders_data)

# boxplot of age (Age.Event) by Type
ggplot(data = leaders_data) +
  geom_boxplot(aes(x = Type, y = Age.Event))

ggplot(data = leaders_data) +
  geom_point(aes(x = Birth.Date, y = Age.Event))

```
# THIS IS OUR MODEL
$$
T_i \sim Weibull(\alpha, \mu) \\
log(\mu) = \beta_0 + \beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}\\
$$
Log of mu is done so that the mu parameter is always positive (the support on mu in Weibull is that mu > 0)

# MORE SPECIFICALLY
$$
T_i \sim Weibull(\alpha, \mu_i) \\
\log(\mu_i) = \beta_0+\beta_1 \text{(Year of Birth_i)}+\beta_2 \text{(Leadership_i = UsPres)} + \beta_3\text{(Leadership_i = ChinaEmp)} \\
+ \beta_4\text{(Leadership_i = DalaiLama)} +\beta_5\text{(Leadership_i = JapanEmp)}
$$
According to Graham, the following is being done by the model above. So basically Stander et al is doing cox proportional, but they just didn't write out this hazard function explicity in the paper.
$$
h_i(t)=h_0(t)\exp(\beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}) \\
$$
hazard of death is a rate -- instantaneous failure rate (probability is integral from 0 to J over it dt)

$$
f(t) = \alpha\lambda t^{\alpha-1}\exp(-\lambda t^\alpha)
$$
run JAGS mcmc on the betas, follow Stander et al code as a guide

```{r}
# filter out the 4 leaders we need to predict

leaders_data <- leaders_data %>%
  mutate(TypeChinaEmp = case_when(Type == "ChinaEmp" ~ 1,
                             TRUE ~ 0),
         TypeDalaiLama = case_when(Type == "DalaiLama" ~ 1,
                                   TRUE ~ 0),
         TypeJapanEmp = case_when(Type == "JapanEmp" ~ 1, 
                                  TRUE ~ 0),
         TypePope = case_when(Type == "Pope" ~ 1, 
                              TRUE ~ 0),
         TypeUsPres = case_when(Type == "UsPres" ~ 1,
                                TRUE ~ 0))

leaders_nopred <- leaders_data %>%
  filter(!(Name %in% c("14th Dalai Lama (Tenzin Gyatso)", "Naruhito (Emperor Reiwa)", "Barack Obama", "Francis")))

leaders_nopred$Birth.Year <- year(leaders_nopred$Birth.Date)

survival_raw <- leaders_nopred$Age.Event
censored <- leaders_nopred$Censored
survival <- ifelse(censored == 0, survival_raw, NA)

censoring_limits <- ifelse(censored == 0, 120, survival_raw)
x_1 <- leaders_nopred$Birth.Year - mean(leaders_nopred$Birth.Year)
x_2 <- leaders_nopred$TypeUsPres
x_3 <- leaders_nopred$TypeChinaEmp
x_4 <- leaders_nopred$TypeDalaiLama
x_5 <- leaders_nopred$TypeJapanEmp
```

```{r Censor-1st-approach-1}
first_approach <- function(){
  for(i in 1:n){
    #
    # Likelihood part
    #
    censored[i] ~ dinterval(survival[i], censoring_limits[i]) 
    # To deal with censoring: 
    # censored equals 1 for Benedict XVI as of 25/12/2016, 
    # 0 for the other popes.
    # survival equals NA for Benedict XVI,
    # and the survival times of the other popes.
    # censoring_limits equals 11.7 for Benedict XVI,
    # and values (e.g. 32) greater than or equal to 
    # the survival times of the other popes.
    survival[i] ~ dweib(r, mu[i]) # Basic Weibull assumption
    mu[i] <- exp(beta[i])  # Defining beta as log(mu)
    beta[i] <- beta_0 + beta_1*x_1[i] + beta_2*x_2[i] + beta_3*x_3[i] + beta_4*x_4[i] + beta_5*x_5[i]
    # beta = log(mu) is a linear function of the covariates
  }
  
  # Priors
  beta_0 ~ dnorm(0.0, 1.0E-4) # Prior on beta_0 is normal with low precision
  beta_1 ~ dnorm(0.0, 1.0E-4) # Prior on beta_1 is normal with low precision
  beta_2 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  beta_3 ~ dnorm(0.0, 1.0E-4) # Prior on beta_1 is normal with low precision
  beta_4 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  beta_5 ~ dnorm(0.0, 1.0E-4) # Prior on beta_2 is normal with low precision
  r ~ dexp(0.001) # Prior on r
  
  # Define the alphas
  alpha_0 <- - beta_0 / r
  alpha_1 <- - beta_1 / r
  alpha_2 <- - beta_2 / r
  alpha_3 <- - beta_3 / r
  alpha_4 <- - beta_4 / r
  alpha_5 <- - beta_5 / r

  # Percentage increases
  # percentage_increase_age <- 100*(exp(alpha_1) - 1)
  # percentage_increase_year <- 100*(exp(alpha_2) - 1)
  
  # Posterior median at various covariate values
  beta_med <- beta_0 + beta_1*x_1_new + beta_2*x_2 + beta_3*x_3 + beta_4*x_4 + beta_5*x_5 
  # New values need to be supplied
  t_median <- pow(log(2) * exp(-beta_med), 1 / r)
  
  # Predictive distribution of age at the new values
  
  beta_Francis <- beta_0 + beta_1*year_Francis 
  # Values of age_Francis and year_Francis need to be provided
  mu_Francis <- exp(beta_Francis)
  survival_Francis ~ dweib(r, mu_Francis) %_% T(present_length, upper_length) 
  # Take into account the current pontificate length
  # Also specify a sensible upper bound
  age_Francis_predictive <- survival_Francis 
  # Work also with age
}
```

```{r 1st-approach-Francis-pred, warning=F, message=F}
library(R2jags)
# Prepare the data
age_new <- 60
x_1_new <-  age_new - mean(leaders_nopred$Birth.Year)
# year_new <- 1750
# x_2_new <- year_new - mean(popes$Year.Elected)

# Data specific to Francis
age_at_election <- 76.29
age_Francis <- age_at_election - mean(popes$Age.Election)
year_Francis <- 2013 - mean(popes$Year.Elected)

# Predictive distribution of pontificate length for Francis at election
# Conditional on being greater than 7.46 years (Aug 24, 2020)
present_length <- 7.46
upper_length <- 25
data_Popes_first_approach <- list("n",
                   "censored",
                   "survival",
                   "censoring_limits",
                   "x_1",
                   "x_2",
                   "x_1_new",
                   #"x_2_new",
                   "age_at_election",
                   "age_Francis",
                   "year_Francis",
                   "present_length",
                   "upper_length")
Bayesian_Popes_first_approach <- jags(data = data_Popes_first_approach,  
                       parameters.to.save = c("beta_0",
                                              "beta_1",
                                              "beta_2",
                                              "alpha_0",
                                              "alpha_1",
                                              "alpha_2",
                                              #"percentage_increase_age",
                                              #"percentage_increase_year",
                                              "r",
                                              "t_median",
                                              "survival_Francis",
                                              "age_Francis_predictive"), 
                       n.iter = 50000, 
                       n.chains = 3,
                       model.file = first_approach)

Bayesian_Popes_first_approach
```





