---
title: "second approach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
#library(R2jags)
library(R2jags)
library(tidyverse)
library(lubridate)
library(knitr)
library(broom)
```

```{r}
leaders_data <- get(load("Leaders.RData"))

leaders_data$Censored <- as.factor(leaders_data$Censored)
leaders_data$Type <- as.factor(leaders_data$Type)
leaders_data$Age.Event <- as.numeric(leaders_data$Age.Event)
summary(leaders_data)
```

# Boxplot of age (Age.Event) by Type
```{r}
ggplot(data = leaders_data) +
  geom_boxplot(aes(x = Type, y = Age.Event))

ggplot(data = leaders_data) +
  geom_point(aes(x = Birth.Date, y = Age.Event))


```

# Our Model
$$
T_i \sim Weibull(\alpha, \mu) \\
log(\mu) = \beta_0 + \beta_1x_{i1}+\beta_2x_{i2}+...+\beta_px_{ip}\\
$$
Log of mu is done so that the mu parameter is always positive (the support on mu in Weibull is that mu > 0)

### More specificically,
$$
T_i \sim Weibull(r, \mu_i) \\
\log(\mu_i) = \beta_0+\beta_1 \text{(Year of Birth_i)}+\beta_2 \text{(Leadership_i = UsPres)} + \beta_3\text{(Leadership_i = ChinaEmp)} \\
+ \beta_4\text{(Leadership_i = DalaiLama)} +\beta_5\text{(Leadership_i = JapanEmp)} \\
+ \beta_6\text{(Year of Birth_i * (Leadership_i = UsPres))} \\
+ \beta_7\text{(Year of Birth_i * (Leadership_i = ChinaEmp))} \\
+ \beta_8\text{(Year of Birth_i * (Leadership_i = DalaiLama))} \\
+ \beta_9\text{(Year of Birth_i * (Leadership_i = JapanEmp))}
$$


# Data Cleaning
```{r}
leaders_data$Birth.Year <- year(leaders_data$Birth.Date)

leaders_data <- leaders_data %>%
  mutate(TypeChinaEmp = as.factor(case_when(Type == "ChinaEmp" ~ 1,
                             TRUE ~ 0)),
         TypeDalaiLama = as.factor(case_when(Type == "DalaiLama" ~ 1,
                                   TRUE ~ 0)),
         TypeJapanEmp = as.factor(case_when(Type == "JapanEmp" ~ 1, 
                                  TRUE ~ 0)),
         TypePope = as.factor(case_when(Type == "Pope" ~ 1, 
                              TRUE ~ 0)),
         TypeUsPres = as.factor(case_when(Type == "UsPres" ~ 1,
                                TRUE ~ 0)))

leaders_nopred <- leaders_data %>%
  filter(!(Name %in% c("14th Dalai Lama (Tenzin Gyatso)", "Naruhito (Emperor Reiwa)", "Barack Obama", "Francis")))

survival_raw <- leaders_nopred$Age.Event
n <- length(survival_raw)

censored <- leaders_nopred$Censored
survival <- ifelse(censored == 0, survival_raw, NA)

censoring_limits <- ifelse(censored == 0, 100, survival_raw)
x_1 <- leaders_nopred$Birth.Year - mean(leaders_nopred$Birth.Year)
x_2 <- leaders_nopred$TypeUsPres
x_3 <- leaders_nopred$TypeChinaEmp
x_4 <- leaders_nopred$TypeDalaiLama
x_5 <- leaders_nopred$TypeJapanEmp
# add interaction b/w Birth.Year and Type
x_6 <- x_1*as.numeric(as.character(x_2))
x_7 <- x_1*as.numeric(as.character(x_3))
x_8 <- x_1*as.numeric(as.character(x_4))
x_9 <- x_1*as.numeric(as.character(x_5))
```


```{r}
set.seed(123)
second_approach <- function() {
  for(i in 1:n_censored) {
    z_censored[i] ~ dpois(phi_censored[i])
    phi_censored[i] <- mu_censored[i] * pow(t_censored[i], r)
    mu_censored[i] <- exp(beta_censored[i])
    beta_censored[i] <- beta_0 + beta_1*x_1_censored[i] + beta_2*x_2_censored[i] + beta_3*x_3_censored[i] + beta_4*x_4_censored[i] + beta_5*x_5_censored[i] + beta_6*x_6_censored[i] + beta_7*x_7_censored[i] + beta_8*x_8_censored[i] + beta_9*x_9_censored[i]
  }
  
  for(j in 1:n_non_censored) { #total rows - 7 ** 7 are censored, rest are not censored
    survival_non_censored[j] ~ dweib(r, mu[j])
    mu[j] <- exp(beta[j])
    beta[j] <- beta_0 + beta_1*x_1_non_censored[j] + beta_2*x_2_non_censored[j] + beta_3*x_3_non_censored[j] + beta_4*x_4_non_censored[j] + beta_5*x_5_non_censored[j] + beta_6*x_6_non_censored[j] + beta_7*x_7_non_censored[j] + beta_8*x_8_non_censored[j] + beta_9*x_9_non_censored[j]
  }
  
  beta_0 ~ dnorm(0.0, 1.0E-3) # Prior on beta_0 is normal with low precision
  beta_1 ~ dnorm(0.0, 1.0E-3) # Prior on beta_1 is normal with low precision
  beta_2 ~ dnorm(0.0, 1.0E-3) # Prior on beta_2 is normal with low precision
  beta_3 ~ dnorm(0.0, 1.0E-3) # Prior on beta_3 is normal with low precision
  beta_4 ~ dnorm(0.0, 1.0E-3) # Prior on beta_4 is normal with low precision
  beta_5 ~ dnorm(0.0, 1.0E-3) # Prior on beta_5 is normal with low precision
  beta_6 ~ dnorm(0.0, 1.0E-3) # Prior on beta_6 is normal with low precision
  beta_7 ~ dnorm(0.0, 1.0E-3) # Prior on beta_7 is normal with low precision
  beta_8 ~ dnorm(0.0, 1.0E-3) # Prior on beta_8 is normal with low precision
  beta_9 ~ dnorm(0.0, 1.0E-3) # Prior on beta_9 is normal with low precision
  r ~ dexp(1) # Prior on r
  
  # define alphas
  alpha_0 <- - beta_0 / r
  alpha_1 <- - beta_1 / r
  alpha_2 <- - beta_2 / r
  alpha_3 <- - beta_3 / r
  alpha_4 <- - beta_4 / r
  alpha_5 <- - beta_5 / r  
  alpha_6 <- - beta_6 / r
  alpha_7 <- - beta_7 / r
  alpha_8 <- - beta_8 / r
  alpha_9 <- - beta_9 / r
  
  # Percentage increases
  percentage_increase_year <- 100*(exp(alpha_1) - 1)
  percentage_increase_UsPres <- 100*(exp(alpha_2) - 1)
  percentage_increase_ChinaEmp <- 100*(exp(alpha_3) - 1)
  percentage_increase_DalaiLama <- 100*(exp(alpha_4) - 1)
  percentage_increase_JapanEmp <- 100*(exp(alpha_5) - 1)
  percentage_increase_BYUsPres <- 100*(exp(alpha_6) - 1)
  percentage_increase_BYChinaEmp <- 100*(exp(alpha_7) - 1)
  percentage_increase_BYDalaiLama <- 100*(exp(alpha_8) - 1)
  percentage_increase_BYJapanEmp <- 100*(exp(alpha_9) - 1)
  
  # Predictive distribution of age at the new values
  beta_Francis <- beta_0 + beta_1*year_Francis
  mu_Francis <- exp(beta_Francis)
  survival_Francis ~ dweib(r, mu_Francis) %_% T(present_length_Francis, upper_length)
  age_Francis_predictive <- survival_Francis

  beta_Obama <- beta_0 + (beta_1+beta_6)*year_Obama + beta_2
  mu_Obama <- exp(beta_Obama)
  survival_Obama ~ dweib(r, mu_Obama) %_% T(present_length_Obama, upper_length)
  age_Obama_predictive <- survival_Obama

  beta_Dalai <- beta_0 + (beta_1+beta_8)*year_Dalai + beta_4
  mu_Dalai <- exp(beta_Dalai)
  survival_Dalai ~ dweib(r, mu_Dalai) %_% T(present_length_Dalai, upper_length)
  age_Dalai_predictive <- survival_Dalai
  
  beta_Naruhito <- beta_0 + (beta_1+beta_9)*year_Naruhito + beta_5
  mu_Naruhito <- exp(beta_Naruhito)
  survival_Naruhito ~ dweib(r, mu_Naruhito) %_% T(present_length_Naruhito, upper_length)
  age_Naruhito_predictive <- survival_Naruhito
}
```


```{r}
censored_dex = which(censored==1) # index of ppl who are censored

z_censored <- rep(0, length(censored_dex))
t_censored <- censoring_limits[censored_dex]
x_1_censored <- x_1[censored_dex] 
x_2_censored <- x_2[censored_dex] 
x_3_censored <- x_3[censored_dex]
x_4_censored <- x_4[censored_dex]
x_5_censored <- x_5[censored_dex]
x_6_censored <- x_6[censored_dex]
x_7_censored <- x_7[censored_dex]
x_8_censored <- x_8[censored_dex]
x_9_censored <- x_9[censored_dex]

n_censored <- length(censored_dex)

survival_non_censored <- survival[-censored_dex] # Remove ALL ALIVE PEOPLE
x_1_non_censored <- x_1[-censored_dex]
x_2_non_censored <- x_2[-censored_dex]
x_3_non_censored <- x_3[-censored_dex]
x_4_non_censored <- x_4[-censored_dex]
x_5_non_censored <- x_5[-censored_dex]
x_6_non_censored <- x_6[-censored_dex]
x_7_non_censored <- x_7[-censored_dex]
x_8_non_censored <- x_8[-censored_dex]
x_9_non_censored <- x_9[-censored_dex]

n_non_censored <- length(survival_non_censored)

birth_year_Francis <- leaders_data$Birth.Year[leaders_data$Name == "Francis"]
year_Francis <- birth_year_Francis - mean(leaders_nopred$Birth.Year)
present_length_Francis <- leaders_data$Age.Event[leaders_data$Name =="Francis"]

birth_year_Obama <- leaders_data$Birth.Year[leaders_data$Name == "Barack Obama"]
year_Obama <- birth_year_Obama - mean(leaders_nopred$Birth.Year)
present_length_Obama <- leaders_data$Age.Event[leaders_data$Name =="Barack Obama"]

birth_year_Dalai <- leaders_data$Birth.Year[leaders_data$Name == "14th Dalai Lama (Tenzin Gyatso)"]
year_Dalai <- birth_year_Dalai - mean(leaders_nopred$Birth.Year)
present_length_Dalai <- leaders_data$Age.Event[leaders_data$Name =="14th Dalai Lama (Tenzin Gyatso)"]

birth_year_Naruhito <- leaders_data$Birth.Year[leaders_data$Name == "Naruhito (Emperor Reiwa)"]
year_Naruhito <- birth_year_Naruhito - mean(leaders_nopred$Birth.Year)
present_length_Naruhito <- leaders_data$Age.Event[leaders_data$Name =="Naruhito (Emperor Reiwa)"]

upper_length <- 100

# REMEBER TO RENAME FUNCTION data_Popes_second_approach

data_Popes_second_approach <- list("n_censored",
                                   "n_non_censored",
                                   "z_censored",
                                   "t_censored",
                                   "x_1_censored",
                                   "x_2_censored",
                                   "x_3_censored",
                                   "x_4_censored",
                                   "x_5_censored",
                                   "x_6_censored",
                                   "x_7_censored",
                                   "x_8_censored",
                                   "x_9_censored",
                                   "survival_non_censored",
                                   "x_1_non_censored",
                                   "x_2_non_censored",
                                   "x_3_non_censored",
                                   "x_4_non_censored",
                                   "x_5_non_censored",
                                   "x_6_non_censored",
                                   "x_7_non_censored",
                                   "x_8_non_censored",
                                   "x_9_non_censored", 
                                   "year_Francis",
                                   "year_Obama", 
                                   "year_Dalai",
                                   "year_Naruhito",
                                   "present_length_Francis", 
                                   "present_length_Obama",
                                   "present_length_Dalai",
                                   "present_length_Naruhito",
                                   "upper_length")

Bayesian_Popes_second_approach <- jags(data = data_Popes_second_approach,  
                                      parameters.to.save = c("beta_0",
                                                             "beta_1",
                                                             "beta_2", 
                                                             "beta_3", 
                                                             "beta_4", 
                                                             "beta_5",
                                                             "beta_6",
                                                             "beta_7", 
                                                             "beta_8", 
                                                             "beta_9",
                                                             "r", 
                                                             "alpha_0",
                                                             "alpha_1",
                                                             "alpha_2",
                                                             "alpha_3",
                                                             "alpha_4",
                                                             "alpha_5",
                                                             "alpha_6",
                                                             "alpha_7",
                                                             "alpha_8",
                                                             "alpha_9",
                                                             "percentage_increase_year",
                                                             "percentage_increase_UsPres",
                                                             "percentage_increase_ChinaEmp",
                                                             "percentage_increase_DalaiLama",
                                                             "percentage_increase_JapanEmp", 
                                                             "percentage_increase_BYUsPres",
                                                             "percentage_increase_BYChinaEmp",
                                                             "percentage_increase_BYDalaiLama",
                                                             "percentage_increase_BYJapanEmp",
                                                             "age_Francis_predictive",
                                                             "age_Obama_predictive",
                                                             "age_Dalai_predictive",
                                                             "age_Naruhito_predictive",
                                                             "survival_non_censored"
                                                             ), 

                                      n.iter = 50000, 
                                      n.chains = 3,
                                      model.file = second_approach)

Bayesian_Popes_second_approach
```

```{r}
library(coda)
parameters = c("beta_0", "beta_1", "beta_2", "beta_3", "beta_4", "beta_5", 
               "beta_6", "beta_7",  "beta_8", "beta_9", "r")

res = data.frame(Bayesian_Popes_second_approach$BUGSoutput$sims.matrix)[,parameters]
colnames(res) <- gsub("^.*\\.","", colnames(res) )

# lapply(seq(1,length(res)), function(i) {
#   traceplot(as.mcmc(res[,i]), smooth = FALSE, type = "l",
#             xlab = "Iterations", ylab = colnames(res)[i],
#             main = paste("Traceplot of ", colnames(res)[i]))
# })

tps <- function(var){
  ggplot(res, aes_(y=as.name(var), x=seq(1,nrow(res)))) +
    geom_line() +
    labs(title=paste("Traceplot of ", as.name(var)),
         x ="Iterations", y = as.name(var))
}
lapply(names(res), tps)
```

```{r}
library(ggplot2)
res_lag1 = lapply(seq(1:length(res)), function(i) {
  lres = lag(res[,i],1)
  plot(y=res[,i], x= lres, 
       xlab = paste0(colnames(res)[i], "-1"),
       ylab = paste0(colnames(res)[i]),
       main = paste("Lag-1 Scatter Plot of", colnames(res)[i]))
})
```

```{r}
lapply(seq(1,length(res)), function(i) { 
  acf(res[,i], xlab = "Lag", ylab = "ACF", 
            main = paste("ACF Plot of ", colnames(res)[i]))
})
```

```{r}
people = c("age_Francis_predictive", "age_Obama_predictive", "age_Dalai_predictive",
"age_Naruhito_predictive")
four_pred = data.frame(Bayesian_Popes_second_approach$BUGSoutput$sims.matrix)[,people]

print(paste("P(Dalai Lifespan > Francis Lifespan)=", mean(four_pred$age_Dalai_predictive>four_pred$age_Francis_predictive)))

print(paste("P(Obama Lifespan > Naruhito Lifespan)=", mean(four_pred$age_Obama_predictive>four_pred$age_Naruhito_predictive)))
```

# Posterior Predictive Plots (4 separate)

```{r}
ppd_plot <- function(i) {
  name = gsub(".*[_]([^.]+)[_].*", "\\1", colnames(four_pred)[i])
  minval = round(min(four_pred[,i]), 2)
  medval = round(median(four_pred[,i]), 2)
  val95 = round(quantile(four_pred[,i], 0.95),2)
  maxval = round(max(four_pred[,i]), 2)
ggplot(four_pred, aes(x=four_pred[,i])) + geom_histogram(aes(y=..density..), 
                                                   binwidth = 0.8,
                                                   boundary = minval) +
  geom_vline(xintercept = minval, colour = "black")+
  geom_vline(xintercept = medval, colour = "black") +
  geom_vline(xintercept = val95, colour = "black")+
  geom_hline(yintercept = 0, colour = "black") +
  xlab("Lifespan (years)") +
  ylab("Predictive Probability density") +
  ggtitle(paste("Predictive Posterior Distribution for Lifespan of", name)) +
  ylim(c(0, 0.07))+
  annotate("text", x = minval+0.15*(medval-minval), y = 0.06, label = paste("Current age", minval, "years", sep = "\n"), size = 3)+
  annotate("text", x = (medval+minval)/2, y = 0.05, label = "50%", size = 3)+
  annotate("text", x = medval+0.15*(val95-medval), y = 0.06, label = paste("Median", medval,"years", sep = "\n"), size = 3)+
  annotate("text", x = (medval+val95)/2, y = 0.05, label = "45%", size = 3)+
  annotate("text", x = val95+2.2, y = 0.06, label = paste("95%", "quantile", val95, "years", sep = "\n"), size = 3)+
  annotate("text", x = val95+3, y = 0.05, label = "5%", size = 3)
}

lapply(seq(1,4), ppd_plot)
```

## Posterior Predictive Plots (2 pairs)

### Francis and Dalai Lama 

```{r}
francis_dalai = data.frame(plifespans = c(four_pred$age_Francis_predictive, four_pred$age_Dalai_predictive), person = c(rep("Francis", nrow(four_pred)), rep("Dalai", nrow(four_pred))))

mins = francis_dalai %>%
  group_by(person) %>%
  summarise(min_vals = round(min(plifespans), 2))

meds = francis_dalai %>%
  group_by(person) %>%
  summarise(med_vals = round(median(plifespans), 2))

val95 = francis_dalai %>%
  group_by(person) %>%
  summarise(vals_95 = round(quantile(plifespans, 0.95), 2))

francis_txt = paste(paste("Current age:", mins$min_vals[mins$person=="Francis"], "yrs"),
      paste("Median:", meds$med_vals[meds$person=="Francis"],"yrs"),
      paste("95% quantile:", val95$vals_95[val95$person=="Francis"], "yrs"),
      sep = "\n")

dalai_txt = paste(paste("Current age:", mins$min_vals[mins$person=="Dalai"], "yrs"),
      paste("Median:",  meds$med_vals[meds$person=="Dalai"],"yrs"),
      paste("95% quantile:", val95$vals_95[val95$person=="Dalai"], "yrs"),
      sep = "\n")

plot_txt <- data.frame(
  label = c(francis_txt, dalai_txt),
  person = c("Francis", "Dalai")
)

ggplot(francis_dalai, aes(x=plifespans)) + 
  geom_histogram(aes(y=..density..), binwidth = 0.8,boundary = round(min(francis_dalai$plifespans), 2)) +
  facet_wrap(~person) +
  geom_vline(data=mins, aes(xintercept=min_vals))+
  geom_vline(data=meds, aes(xintercept=med_vals))+
  geom_vline(data=val95, aes(xintercept=vals_95))+
  geom_hline(yintercept = 0, colour = "black") +
  xlab("Lifespan (yrs)") + ylab("Predictive Probability density") +
  ggtitle(paste("Posterior Predictive Lifespans (Pope Francis and 14th Dalai Lama)")) +
  ylim(c(0, 0.05)) + 
  geom_text(data = plot_txt, mapping = aes(x = 92.8, y = 0.048, label = label), size = 2.8)
```

### Obama and Naruhito

```{r}
obama_naru = data.frame(plifespans = c(four_pred$age_Obama_predictive, four_pred$age_Naruhito_predictive), person = c(rep("Obama", nrow(four_pred)), rep("Naruhito", nrow(four_pred))))

mins = obama_naru %>%
  group_by(person) %>%
  summarise(min_vals = round(min(plifespans), 2))

meds = obama_naru %>%
  group_by(person) %>%
  summarise(med_vals = round(median(plifespans), 2))

val95 = obama_naru %>%
  group_by(person) %>%
  summarise(vals_95 = round(quantile(plifespans, 0.95), 2))

obama_txt = paste(paste("Current age:", mins$min_vals[mins$person=="Obama"], "yrs"),
      paste("Median:", meds$med_vals[meds$person=="Obama"],"yrs"),
      paste("95% quantile:", val95$vals_95[val95$person=="Obama"], "yrs"),
      sep = "\n")

naru_txt = paste(paste("Current age:", mins$min_vals[mins$person=="Naruhito"], "yrs"),
      paste("Median:",  meds$med_vals[meds$person=="Naruhito"],"yrs"),
      paste("95% quantile:", val95$vals_95[val95$person=="Naruhito"], "yrs"),
      sep = "\n")

plot_txt <- data.frame(
  label = c(obama_txt, naru_txt),
  person = c("Obama", "Naruhito")
)

ggplot(obama_naru, aes(x=plifespans)) + 
  geom_histogram(aes(y=..density..), binwidth = 0.8,boundary = round(min(obama_naru$plifespans), 2)) +
  facet_wrap(~person) +
  geom_vline(data=mins, aes(xintercept=min_vals))+
  geom_vline(data=meds, aes(xintercept=med_vals))+
  geom_vline(data=val95, aes(xintercept=vals_95))+
  geom_hline(yintercept = 0, colour = "black") +
  xlab("Lifespan (yrs)") + ylab("Predictive Probability density") +
  ggtitle(paste("Posterior Predictive Lifespans (President Obama and Emperor Naruhito)")) +
  ylim(c(0, 0.05)) + 
  geom_text(data = plot_txt, mapping = aes(x = 80, y = 0.04, label = label), size = 3)
```

```{r tidy-model-table}
vars = data.frame(Bayesian_Popes_second_approach$BUGSoutput$root.short)
output <- head(data.frame(Bayesian_Popes_second_approach$BUGSoutput$summary), 16)
output <- output %>%
  mutate("Variable" = head(vars,16),
         "Mean" = output$mean,
         "Standard Deviation" = output$sd,
         "2.5% Quantile" = output$X2.5.,
         "Median" = output$X50.,
         "97.5% Quantile" = output$X97.5.
         ) %>%
  select("Variable", "Mean", "Standard Deviation", "2.5% Quantile", "Median", "97.5% Quantile")
coef <- output %>%
  slice(5:16)
kable(coef)
```

```{r standardized-resid}
beta_0_est <- coef[1,2]
beta_1_est <- coef[2,2]
beta_2_est <- coef[3,2]
beta_3_est <- coef[4,2]
beta_4_est <- coef[5,2]
beta_5_est <- coef[6,2]
beta_6_est <- coef[7,2]
beta_7_est <- coef[8,2]
beta_8_est <- coef[9,2]
beta_9_est <- coef[10,2]

no_living <- leaders_data %>%
  filter(Censored == 0)

estimates <- no_living %>%
  mutate(Birth.Year.Cent = Birth.Year - mean(leaders_nopred$Birth.Year)) %>%
  mutate(ChinaEmp =case_when(Type == "ChinaEmp" ~ 1, TRUE ~ 0),
         DalaiLama =case_when(Type == "DalaiLama" ~ 1, TRUE ~ 0),
         JapanEmp =case_when(Type == "JapanEmp" ~ 1, TRUE ~ 0),
         Pope =case_when(Type == "Pope" ~ 1, TRUE ~ 0),
         UsPres =case_when(Type == "UsPres" ~ 1, TRUE ~ 0)) %>%
  mutate(BYUsPres = Birth.Year.Cent*UsPres) %>%
  mutate(BYChinaEmp = Birth.Year.Cent*ChinaEmp) %>%
  mutate(BYDalaiLama = Birth.Year.Cent*DalaiLama) %>%
  mutate(BYJapanEmp = Birth.Year.Cent*JapanEmp) %>%
  mutate(intercept = 1) %>%
  mutate(predicted = exp(beta_0_est + beta_1_est*(Birth.Year.Cent) + beta_2_est*(UsPres) + beta_3_est*(ChinaEmp) + beta_4_est*(DalaiLama) + beta_5_est*(JapanEmp) + beta_6_est*(UsPres)*(Birth.Year.Cent) + beta_7_est*(ChinaEmp)*(Birth.Year.Cent) + beta_8_est*(DalaiLama)*(Birth.Year.Cent) + beta_9_est*(JapanEmp)*(Birth.Year.Cent)))


# check using matrices -- Ts in the 200s
data.mat = as.matrix(estimates[,c("intercept", "Birth.Year.Cent", "UsPres", "ChinaEmp", "DalaiLama", "JapanEmp","BYUsPres", "BYChinaEmp", "BYDalaiLama", "BYJapanEmp")], nrow = nrow(estimates), byrow = TRUE)

r = Bayesian_Popes_second_approach$BUGSoutput$summary[,1][16]
betas = Bayesian_Popes_second_approach$BUGSoutput$summary[,1][5:14]
beta.mat = as.matrix(betas, nrow = p, byrow = TRUE)
logmus = data.mat %*% beta.mat

min.age = min(leaders_nopred$Age.Event)

install.packages("TruncatedDistributions", repos="http://R-Forge.R-project.org")
library(TruncatedDistributions)
set.seed(123)
Ts = rtweibull(166, r, (1/exp(logmus))^(1/r), 0, 100) 

resid = estimates$Age.Event - Ts
resid * estimates$predicted

meanTs = mean(Ts)
diff = Ts - meanTs
sd = sqrt((1/166)*(sum((diff)^2)) * (1+ (1/166) + (diff)^2/sum((diff)^2)))

ggplot(data = estimates, aes(x = Ts, y = resid/sd)) + geom_point()
ggplot(data = estimates, aes(x = Birth.Year, y = resid/sd)) + geom_point()
```
```

## Posterior Predictive Checks

```{r}
non_censored_prediction <- filter(leaders_data, Censored == 0)
K <- 10
ysims <- matrix(nrow = nrow(res), ncol = K)
for(k in 1:K){
  for(i in 1:nrow(non_censored_prediction)){
    w <- sample(nrow(res), 1)
    samp <- res[w,]
    val <- non_censored_prediction[i, ]
    beta_temp <- samp$beta_0 + samp$beta_1*val$Birth.Year + samp$beta_2*as.integer(val$TypeUsPres) + samp$beta_3*as.integer(val$TypeChinaEmp) + samp$beta_4*as.integer(val$TypeDalaiLama) + samp$beta_5*as.integer(val$TypeJapanEmp) + samp$beta_6*val$Birth.Year*as.integer(val$TypeUsPres) + samp$beta_7*val$Birth.Year*as.integer(val$TypeChinaEmp) + samp$beta_8*val$Birth.Year*as.integer(val$TypeDalaiLama) + samp$beta_9*val$Birth.Year*as.integer(val$TypeJapanEmp)
    mu <- exp(beta_temp)
    t <- dweibull(samp$r, mu)
    ysims[i, k] <- t
  }
}
```

```{r}
r <- sample(nrow(non_censored_prediction), 1)
hist(ysims[r,])
abline(v=non_censored_prediction$Age.Event[r], col = "red")
print(mean(ysims<0))
print(mean(non_censored_prediction$Age.Event <0))
```

## Lexis Diagram
```{r}
#lg <- lexis_grid(1960, 2000, 30, 80, delta=5) 
#leaders_data$Initial.Age <- year(leaders_data$Event.Date) - leaders_data$Birth.Year
#subset <- filter(leaders_data, Censored == 0)
#subset <- subset[0:5,]
#ggplot() + geom_polygon(data = subset, mapping = aes(x = Birth.Year, y = Age.Event, group = Name)) + geom_point(data = subset, aes(x = Birth.Year, y = Birth.Year + Age.Event)) + geom_polygon(data = subset, mapping = aes(x = Birth.Year, y = ))
```

```{r}
subset <- filter(leaders_data, Censored == 0)
ggplot(data = subset, aes(x = Birth.Year, Age.Event)) + geom_point(color="pink") + ggtitle("Distribution of Leaders' Ages") + labs(x = "Birth Year", y = "Age")
```